Selection of Polynomial Features by Genetic Algorithms

### 2013-07-22 texts/paper-draft.tex
it works….

\documentclass[12pt]{article}
\usepackage[latin1]{inputenc}
\usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
\geometry{a4paper} % or letter or a5paper or ... etc
% \geometry{landscape} % rotated page geometry
\usepackage{amsfonts}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}

\usepackage{graphicx}
\usepackage{acronym}
\usepackage{pgfplots}
\pgfplotsset{compat=1.8}
\usepgfplotslibrary{statistics}

\usepackage{algpseudocode}
\usepackage{algorithm}

\title{Selection of Polynomial Features by Genetic Algorithms}
\author{Francisco Coelho \and João Neto}
\date{[{\sc draft} \today]} % delete this line to display the current date

\newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}

\acrodef{GA}{Genetic algorithms}
\acrodef{rmse}{root-mean-square error}
\acrodef{GAPoly}[{\sc GAPoly}]{Genetic Algorithms for Polynomials}

%%% BEGIN DOCUMENT
\begin{document}

@@

\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application. If a nonlinear (or otherwise) model requires parameterization, \acp{GA} provide a simple and often effective approach to search for locally optimal parameters. Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \cite{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings. Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \cite{holland1975adaptation}.

One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions. This syntatic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, the subarea of machine learning actually producing the most promising results (\emph{e.g.} \cite{tarlowstochastic}). It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.

@@

This section is dedicated to describe an algorithm to find a polynomial regression from a given dataset. It starts with a brief introduction and description of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.

A usual representation of polynomials is
$$
p\left( x_1, \ldots, x_k\right) = \sum_i \theta_i m_i
$$
where each $m_i$ is a monomial, $m_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$. For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$ coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$. The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$. For the example above
$$
A = \left[ 
\begin{array}{ccc}
1 & 0 & 0 \\
0 & 1 & 1 \\
2 & 0 & 1
\end{array}
\right] \sim 
\left[ 
\begin{array}{ccc}
x_1 \\
x_2 x_3 \\
x_1^2 x_3
\end{array}
\right]
$$
where each row defines a monomial and each column represents a variable. Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
%Notice that if we interpret matrix $A$ as a single column of monomials and set the coefficients $T = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
%$$
%p\left( x_1, x_2, x_3 \right) = TA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
%\begin{array}{ccc}
%x_1 \\
%x_2 x_3 \\
%x_1^2 x_3
%\end{array}
%\right].
%$$

@@

The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials. For each polynomial $p\in \mathcal{P}$ let $\mathcal{M}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
%
\begin{eqnarray*}
\phi_p &=& \min_\Theta J\left(\Theta;\mathcal{M}_p,\mathcal{D}\right)
\end{eqnarray*}
%
by solving the first problem. With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $\mathcal{P}$ until a reasonable approximation of a local minimum is found. 
%
\begin{algorithm}[htb]
\begin{algorithmic}
\Function{GAPoly}{$D,pop_0,\epsilon$}
	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
	\While{$err > \epsilon$}
		\State $pop \gets \Call{IterateGA}{pop}$\Comment{Build next generation}
		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$\Comment{Sort by \ac{rmse} of linear regression}
		\State $err \gets \textsc{rmse}\left(\Call{First}{pop}\right)$
	\EndWhile
	\State\Return $\Call{First}{pop}$
\EndFunction
\end{algorithmic}
\caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials. At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
\end{algorithm} 
%
\begin{theorem}
The composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the \ac{rmse} function, encapsulated in the fitness function $J$.
\end{theorem}
%
\begin{proof}
It follows from the properties of \acp{GA} and linear regression.
\end{proof}
%
Subsection \ref{subs:polynomial.encoding} details of the encoding of individual polynomial instances as chromosomes and other parameters of the \ac{GA} implementation used. The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.
\subsection{Polinomial Encoding}\label{subs:polynomial.encoding}

The encoding for any polynomial will be as follows:
\begin{enumerate}
\item an initial segment detailing which monomials are active (the 1st monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
\item the remaining is split into $m$ sets of bits of equal size, each one representing a monomial
\item each monomial is split into $n$ sets of $d$ size each, i.e., a variable
\item for each variable, the remaining bits give the binary description of the variable degree, i.e., the maximum exponent is given by $2^d-1$
\end{enumerate}

@@

This fitness function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities. One way to balance this is to provide a regularization term into the fitness function. Our proposal is to include a multiplicative factor into the fitness function proportional to the number of monomials. Thus, the fitness function becomes,

$$
J_{fit}\left( \Theta; D \right) = \lambda^{m} \sqrt{ \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2 }
$$

where $m$ is the number of monomials the polynomial has. A $\lambda$ greater than one punishes polynomials with more monomials.

@@

\begin{figure}[htb]\begin{center}
\begin{tikzpicture}
\begin{axis}[width=\textwidth,height=6cm,
	boxplot/draw direction=y,
	xtick={1,2,3,4,5,6,7,8,9,10,11,12,13},
	xticklabels={0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30},
	xlabel={regularisation exponent ($\lambda$)},
	ymin=2.0,
	ymax=3.0,
	ylabel={Error distribution (\ac{rmse})},
	title={Error distribution by regularisation exponent (\textbf{Abalone dataset})},
	mark options={scale=0.5},
	]
	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/figure1.dat};		
	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=7+0.2*rand,y index=6] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=8+0.2*rand,y index=7] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=9+0.2*rand,y index=8] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=10+0.2*rand,y index=9] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=11+0.2*rand,y index=10] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=12+0.2*rand,y index=11] {data/figure1.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=13+0.2*rand,y index=12] {data/figure1.dat};

	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
	\addplot[boxplot] table [y index=0] {data/figure1.dat};
	\addplot[boxplot] table [y index=1] {data/figure1.dat};
	\addplot[boxplot] table [y index=2] {data/figure1.dat};
	\addplot[boxplot] table [y index=3] {data/figure1.dat};
	\addplot[boxplot] table [y index=4] {data/figure1.dat};
	\addplot[boxplot] table [y index=5] {data/figure1.dat};
	\addplot[boxplot] table [y index=6] {data/figure1.dat};
	\addplot[boxplot] table [y index=7] {data/figure1.dat};		
	\addplot[boxplot] table [y index=8] {data/figure1.dat};
	\addplot[boxplot] table [y index=9] {data/figure1.dat};
	\addplot[boxplot] table [y index=10] {data/figure1.dat};
	\addplot[boxplot] table [y index=11] {data/figure1.dat};
	\addplot[boxplot] table [y index=12] {data/figure1.dat};

@@

@@

\begin{figure}[htb]\begin{center}
\begin{tikzpicture}
\begin{axis}[width=\textwidth,height=6cm,
	boxplot/draw direction=y,
	xtick={1,2,3,4,5,6,7,8,9,10,11,12,13},
	xticklabels={0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30},
	xlabel={regularisation exponent ($\lambda$)},
	ymin=2.0,
	ymax=5.0,
	ylabel={Error distribution (\ac{rmse})},
	title={Error distribution by regularisation exponent (\textbf{Auto-MPG dataset})},
	mark options={scale=0.5},
	]
	\addplot[only marks, mark=o,color=black] table [x expr=1+0.2*rand,y index=0] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=2+0.2*rand,y index=1] {data/lambda.errors.Auto-Mpg.dat};		
	\addplot[only marks, mark=o,color=black] table [x expr=3+0.2*rand,y index=2] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=4+0.2*rand,y index=3] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=5+0.2*rand,y index=4] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=6+0.2*rand,y index=5] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=7+0.2*rand,y index=6] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=8+0.2*rand,y index=7] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=9+0.2*rand,y index=8] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=10+0.2*rand,y index=9] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=11+0.2*rand,y index=10] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=12+0.2*rand,y index=11] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[only marks, mark=o,color=black] table [x expr=13+0.2*rand,y index=12] {data/lambda.errors.Auto-Mpg.dat};

	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=1] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=3] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=5] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=7] {data/lambda.errors.Auto-Mpg.dat};		
	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=9] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=11] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};

@@

\begin{figure}[htb]
	\begin{center}
		\includegraphics[height=6cm,width=10cm,angle=0]
			{figures/Abalone_fitnessProgress.eps}
		\caption{fitness progress for Abalone dataset}
		\label{Abalone_fitnessProgress}
	\end{center}
\end{figure}

>>>>

\documentclass[12pt]{article}

%
%	FILE ENCODING
%
\usepackage[latin1]{inputenc}
%

%
%	PAGE GEOMETRY
%
\usepackage{geometry}
\geometry{a4paper}
%

%
%	AMS STUFF
%
\usepackage{amsfonts}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
%

%
%	ACRONYMS
%
\usepackage{acronym}
\acrodef{GA}{Genetic algorithms}
\acrodef{rmse}[error]{root-mean-square error}
\acrodef{GAPoly}[{\sc GAPoly}]{Genetic Algorithms for Polynomials}
%

%
%	FIGURE AND STUFF ENVIRONMENT
%
\usepackage{graphicx}
%

%
%	TIKZ AND RELATED STUFF
%
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{statistics}
%
%	avoid processing figures every time
\usepgfplotslibrary{external}
\tikzexternalize
% ENABLE NEXT LINE TO TYPESET FIGURES
\tikzset{external/force remake}
%

%
%	ALGORITHMS PACKAGES
%
\usepackage{algpseudocode}
\usepackage{algorithm}
%

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

\title{Selection of Polynomial Features by Genetic Algorithms}
\author{Francisco Coelho \and João Neto}
\date{[{\sc draft} \today]}

%
%	DOCUMENT COMMANDS
%
\newcommand{\sample}[1]{\ensuremath{^{\left(#1\right)}}}
%

%
%
%%% BEGIN DOCUMENT
%
%
\begin{document}

@@

\ac{GA} where, arguably, one of the hottest topics of research in the recent decades and with good reason since they outline an optimization scheme easy to conceptualize and with very broad application. If a nonlinear (or otherwise) model requires parameterization \acp{GA} provide a simple and often effective approach to search for locally optimal parameters. Research related to genetic algorithms abound and spans from the 1950s seminal work of Nils Aall Barricelli \cite{barricelli1962numerical} in the Institute for Advanced Study of Princeton to today's principal area of study for thousands of researchers, covered in hundreds of conferences, workshops and other meetings. Perhaps the key impulse to \acp{GA} come from John Holland's work and his book ``Adaptation in Natural and Artificial Systems'' \cite{holland1975adaptation}.

One interesting take on genetic algorithms, named \emph{genetic programming} by John Koza \cite{koza1992genetic}, proposed the use of \acp{GA} to search the syntactic structure of complex functions. This syntatic structure search is keen to the central ideas of deep learning \cite{bengio2013representation, bengio2009learning}, a subarea of machine learning actually producing quite promising results (\emph{e.g.} \cite{tarlowstochastic}). It is also related to the work presented in this paper in the sense that, unlike linear models that have a simple structure, $y=\sum_i \beta_i x_i$, nonlinear (in particular polynomial) models pose an additional ``structure" search problem.

@@

This section is dedicated to the description of an algorithm to find a polynomial regression from a given dataset. It starts with a brief introduction and outline of the algorithm and proceeds into core details as the encoding used to represent individual polynomial instances in the \ac{GA} populations and the regularization of the cost function.

A usual representation of polynomials is
$$
p\left( x_1, \ldots, x_k\right) = \sum_i \theta_i m_i
$$
where each $m_i$ is a monomial, $m_i = \prod_{j} x_j^{\alpha_{ij}}$, the exponents are non-negative integers, $\alpha_{ij}\in\mathbb{N}_0$, and the coefficients are real valued, $\theta_i \in \mathbb{R}$. For example $p\left( x_1, x_2, x_3 \right) = 2 x_1 + x_2 x_3 + \frac{1}{2} x_1^2 x_3$ has monomials $m_1 = x_1, m_2 = x_2 x_3$ and $m_3 = x_1^2 x_3$,    coefficients $\theta_1 = 2, \theta_2 = 1$ and $\theta_3 = 1/2$ and exponents $\alpha_{1,1} = 1, \alpha_{2,2}=1, \alpha_{2,3} = 1, \alpha_{3,1} = 2, \alpha_{3,3} = 1$ and all other $\alpha_{ij} = 0$. The exponents alone are a matrix that defines the monomial structure of the polynomial, $A = \left[\alpha_{ij}\right]$. For the example above
$$
A = \left[ 
\begin{array}{ccc}
1 & 0 & 0 \\
0 & 1 & 1 \\
2 & 0 & 1
\end{array}
\right] \sim 
\left[ 
\begin{array}{ccc}
x_1 \\
x_2 x_3 \\
x_1^2 x_3
\end{array}
\right]
$$
where each row defines a monomial and each column represents a variable. Changing the order of the rows doesn't change the polynomial whereas changing the order of the columns corresponds to changing the respective variables.
%Notice that if we interpret matrix $A$ as a single column of monomials and set the coefficients $T = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]$ then
%$$
%p\left( x_1, x_2, x_3 \right) = TA = \left[ \begin{array}{ccc} 2 & 1 & \frac{1}{2} \end{array} \right ]\left[ 
%\begin{array}{ccc}
%x_1 \\
%x_2 x_3 \\
%x_1^2 x_3
%\end{array}
%\right].
%$$

@@

The second problem is treated in the \ac{GA} setting: Let $\mathcal{D}$ be a dataset as above and $\mathcal{P}$ a set of polynomials. For each polynomial $p\in \mathcal{P}$ let $\mathcal{M}_p$ be the set of monomials in $p$ (without the coefficients) and compute the fitness
%
\begin{eqnarray*}
\phi_p &=& \min_\Theta J\left(\Theta;\mathcal{M}_p,\mathcal{D}\right)
\end{eqnarray*}
%
by solving the first problem. With a fitness of every instance, a \ac{GA} will apply genetic operators (usually mutation and crossover) to evolve the population $\mathcal{P}$ until a reasonable approximation of a local minimum is found. 
%
\begin{algorithm}[tb]
\begin{algorithmic}
\Function{GAPoly}{$D,pop_0,\epsilon$}
	\State $pop \gets pop_0;\: err \gets 1.0+\epsilon$
	\While{$err > \epsilon$}
		\State $pop \gets \Call{IterateGA}{pop}$\Comment{Build next generation}
		\State $pop \gets \Call{Sort}{pop, key = \textsc{Lm.rmse}}$\Comment{Sort by \ac{rmse} of linear regression}
		\State $err \gets \textsc{rmse}\left(\Call{First}{pop}\right)$
	\EndWhile
	\State\Return $\Call{First}{pop}$
\EndFunction
\end{algorithmic}
\caption{\ac{GAPoly} uses linear regression to find monomial coefficients that minimize the \ac{rmse} over a dataset and \acp{GA} to explore the space of polynomials. At exit the \ac{rmse} of the fittest instance is bounded by $\epsilon$.}\label{alg:gapoly}
\end{algorithm} 
%
Notice that the properties of \acp{GA} and linear regression entail that
the composition of \acp{GA} with linear regression, as defined in Algorithm \ref{alg:gapoly}, converges to a polynomial that is a local minimum of the \ac{rmse} function, encapsulated in the fitness function $J$.
%

%
Subsection \ref{subs:polynomial.encoding} details of the encoding of individual polynomial instances as chromosomes and other parameters of the \ac{GA} implementation used. The regularization of the cost function is discussed in subsection \ref{subs:cost.function}.

%
\subsection{Polinomial Encoding}\label{subs:polynomial.encoding}
%
The encoding for any polynomial will be as follows:
\begin{enumerate}
\item an initial segment detailing which monomials are active (the 1st monomial is always active), this is represented in unary description, i.e., each monomial is identified by a single bit
\item the remaining is split into $m$ sets of bits of equal size, each one representing a monomial
\item each monomial is split into $n$ sets of $d$ size each, i.e., a variable
\item for each variable, the remaining bits give the binary description of the variable degree, i.e., the maximum exponent is given by $2^d-1$
\end{enumerate}

@@

This fitness function tends to prefer more complex polynomials, namely in the number of monomials which provides the regression algorithm for more fitting possibilities. One way to balance this is to provide a regularization term into the fitness function. Our proposal is to include a multiplicative factor into the fitness function proportional to the number of monomials. Thus, the fitness function from equation \ref{eq:rmse} becomes
\begin{eqnarray}
J_{fit}\left( \Theta; D \right) &=& \lambda^{m} \sqrt{ \frac{1}{n} \sum_{i=1}^n \left( y^{(i)} - h_{\Theta}\left(x_1^{(i)}, \ldots, x_m^{(i)}\right) \right)^2 }
\label{eq:rmse-reg}
\end{eqnarray}
%
where $m$ is the number of monomials the polynomial has. A $\lambda$ greater than one punishes polynomials with more monomials.

@@

\begin{figure}[tb]\begin{center}
\begin{tabular}{cc}
\begin{tikzpicture}
\begin{axis}[
	width=0.5\textwidth,height=6cm,
	boxplot/draw direction=x,
	ytick={1,2,3,4,5,6,7},
	yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
	title = {Abalone Dataset},
	xlabel = {Error (Abalone)},
	ylabel={Regularization ($\lambda$)},
	]	
	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
	\addplot[boxplot] table [y index=0] {data/figure1.dat};
	\addplot[boxplot] table [y index=2] {data/figure1.dat};
	\addplot[boxplot] table [y index=4] {data/figure1.dat};
	\addplot[boxplot] table [y index=6] {data/figure1.dat};		
	\addplot[boxplot] table [y index=8] {data/figure1.dat};
	\addplot[boxplot] table [y index=10] {data/figure1.dat};
	\addplot[boxplot] table [y index=12] {data/figure1.dat};
\end{axis}	
\end{tikzpicture}
%
% RIGHT PANE
&
%
%
\begin{tikzpicture}
\begin{axis}[
	width=0.5\textwidth,height=6cm,
	boxplot/draw direction=x,
	ytick={1,2,3,4,5,6,7},
	yticklabels={0.7, 0.8, 0.9, 1.0, 1.1, 1.2, 1.3},
	title = {Auto MPG Dataset},
	xlabel={Error (Auto MPG)},
	]	
	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};		
	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};

@@

\end{tabular}

@@

\end{center}\end{figure}
%
%
\end{document}

\begin{figure}[tb]\begin{center}
\begin{tikzpicture}
\begin{axis}[width=\textwidth,height=6cm,
	boxplot/draw direction=y,
	xtick={1,2,3,4,5,6,7,8,9,10,11,12,13},
	xticklabels={0.70, 0.75, 0.80, 0.85, 0.90, 0.95, 1.00, 1.05, 1.10, 1.15, 1.20, 1.25, 1.30},
	xlabel={regularisation exponent ($\lambda$)},
	ymin=2.0,
	ymax=5.0,
	ylabel={Error distribution (\ac{rmse})},
	title={Error distribution by regularisation exponent (\textbf{Auto-MPG dataset})},
	mark options={scale=0.5},
	]	
	\pgfplotsset{boxplot/every box/.append style={fill=gray!50!white}}
	\addplot[boxplot] table [y index=0] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=1] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=2] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=3] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=4] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=5] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=6] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=7] {data/lambda.errors.Auto-Mpg.dat};		
	\addplot[boxplot] table [y index=8] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=9] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=10] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=11] {data/lambda.errors.Auto-Mpg.dat};
	\addplot[boxplot] table [y index=12] {data/lambda.errors.Auto-Mpg.dat};

@@

\begin{figure}[htb]
\begin{center}
\begin{tikzpicture}

\begin{axis}[
	width=0.95\textwidth,
	height=6cm,
	xmin=-5,
	xmax=255,
	mark options={scale=0.5},
	line width=1pt,
	xlabel={number of iterations},
	ylabel={\ac{rmse}},
	title={Error (\ac{rmse}) vs. number of iterations (\textbf{Abalone dataset})},
	]

	\addplot[solid,color=black] table [x expr=\coordindex, y index = 0] {data/lambdas.dat};\addlegendentry{$\lambda=0.975$};
	\addplot[dashed,color=black] table [x expr=\coordindex, y index = 1] {data/lambdas.dat};\addlegendentry{$\lambda=1.000$};
\end{axis}
\end{tikzpicture}
\caption{fitness progress for Abalone dataset (population $200$)}
\label{Abalone_fitnessProgress}

\end{center}
\end{figure}
